Probabilistic Inference of Basic Stellar Parameters:\\ Application to Flickering Stars \altaffilmark{\titledag}

### 2015-06-05 version1.0/HBM.tex
adding words

In order to account for the two-dimensional observational uncertainties in
this data set, we marginalize over the latent, `true' values of flicker,
stellar density and $\log(g)$
\citep[see, e.g.][]{hogg:2010, foreman-mackey:2014, rogers:2015, angus:2015}.
The observed values of the variables, flicker, stellar density and $\log(g)$
can be thought of as a single `draw' from an underlying (assumed Gaussian)
probability distribution, with mean equal to the `true' value and standard
deviation equal to the quoted uncertainty.
We compute the likelihood of the data given the model, marginalized over the
`true' values of the variables, using importance sampling.
We sample from the PDF of the `truth', given the observed value and compute
the likelihood of each of those samples.
This PDF is actually a {\it posterior} PDF, produced by the astronomers who
provide the catalog.
For example, the $log(g)$ values used here are point estimates of the
posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
{\it Kepler} light curves and stellar models.
We would use the posterior samples generated in previous model fitting process
if they were available, however generating new samples is an acceptable
approximation provided the posteriors are Gaussian (or log-Normal) and the
priors used by the previous fitters were uninformative.
Such an approximation {\it will not be necessary} for those studies using
the stellar densities or $\log(g)$s calculated using newly-calibrated relations
presented here, as long as the posterior PDF samples are published!
After producing new posterior PDF samples we then add up the individual log-likelihoods of each sample to compute the total marginalized likelihood for a star.
The marginalized likelihood of the whole dataset is then computed as the sum
of the log-likelihoods of each star.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
TBD
}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
TBD
}
\label{fig:logg}
\end{center}
\end{figure}

>>>>

In order to account for the two-dimensional observational uncertainties in
this data set, we marginalize over the latent, `true' values of flicker,
stellar density and $\log(g)$
\citep[see, e.g.][]{hogg:2010, foreman-mackey:2014, rogers:2015, angus:2015}.
The observed values of the variables; flicker, stellar density and $\log(g)$
can be thought of as a single `draw' from an underlying (assumed Gaussian)
probability distribution, with mean equal to the `true' value and standard
deviation equal to the observational uncertainty.
We compute the likelihood of the data given the model, marginalized over the
`true' values of the variables, using importance sampling.
We sample from the posterior PDFs of the `true', latent variable values,
conditioned on the observed values and compute the likelihood of each of those
samples.
% These posterior PDFs could be produced by the astronomers who provide the
% catalog.
% If posterior samples are made available, they can be used in this stage of
% the inference.
% In the event that they are not provided however, they can be generated.
% For example, the $log(g)$ values used here are point estimates of the
% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
% {\it Kepler} light curves and stellar models.
We would use the posterior samples generated in previous model fitting process
if they were available, however generating new samples is an acceptable
approximation provided the posteriors are Gaussian (or log-Normal) and the
priors used by the previous fitters were uninformative.
Such an approximation {\it will not be necessary} for those studies using
the stellar densities or $\log(g)$s calculated using newly-calibrated relations
presented here, as long as the posterior PDF samples are published!
After producing new posterior PDF samples we then add up the individual log-likelihoods of each sample to compute the total marginalized likelihood for a star.
The marginalized likelihood of the whole dataset is then computed as the sum
of the log-likelihoods of each star.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
Stellar density vs. flicker. This figure shows the model, conditioned on the
data. The solid black line shows the median model values, the dark blue
region shows the width of the intrinsic scatter in the relation; the $\sigma$
paramete. The faint blue region shows twice the width (the 2$\sigma$ region).
}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
the model, conditioned on the data. The solid black line shows the median model
values, the dark pink region shows the width of the intrinsic scatter in the
relation; the $\sigma$ parameter. The faint pink region shows twice the width
(the 2$\sigma$ region).
}
\label{fig:logg}
\end{center}
\end{figure}

### 2015-06-05 version1.0/introduction.tex
adding words

Accurate stellar characterization plays a vital role for many active research 
fields within astronomy. For example, stellar populations, galactic archaeology, 
the study of binary stars, asteroseismology and exoplanet studies all rely on 
inferences of basic stellar parameters to varying degrees. Empirically-derived
and reliable estimates are of particular value, increasing our confidence in the
end-product results built upon these inputs.

Basic stellar parameters, such as effective temperature and surface gravity, 
can be inferred using one (or more) of several types of observations, e.g. 
spectroscopy, photometry, interferometry, etc. This inference can be done by
invoking theoretical models or by building an empirical calibration library.
As an example, an observed stellar spectrum could be matched against a library of 
theoretical spectra generated using stellar atmosphere models, or, against a 
library of observed spectra of ``standard stars'', serving as calibrators.
%Fundamental stellar parameters, such as mass and radius, which cannot be 
%directly inferred from the available observations can often be inferred by 
%invoking theoretical stellar models, e.g. fitting model isochrones to 
%spectroscopic measurements.
Regardless of the approach, be it theoretical or empirical, the methods used
for the inference of stellar parameters are traditionally ``deterministic''.
In this context, a deterministic model can be loosely described as one where
a particular observational input always returns a single-valued output for a 
parameter of interest. In this framework, nature itself has no variance and
our underlying model is considered to be a perfect description of reality.

An alternative approach for inferring model parameters is the 
``probabilistic'' one. Unlike the deterministic case, a single observational 
input is intepretted to be caused by a range of possible model parameters
described by a probability distribution. This statement is true even in the case
of a perfect observation of infinite signal-to-noise, since our underlying 
model itself comes with uncertainty. In recent years, there has been a shift 
towards such methods in several areas of astronomy, particularly within the 
exoplanet community. For example, \citet{wolfgang:2015} considered that
the mass-radius relationship of exoplanets is probabilistic, since a particular
sized planet could be due to a range of planet masses, thanks to variances in
compositions, environment and other circumstances. These recent demonstrations
in exoplanetary science have prompted us to consider the need for treating the 
parent stars in the same probabilistic framework, with potential applications 
spanning many fields of astronomy.

The demand for probilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, but also... (ruth will continue this sentence)

>>>>

Accurate stellar characterization plays a vital role for many active research
fields within astronomy.
For example, stellar populations, galactic archaeology, the study of binary
stars, asteroseismology and exoplanet studies all rely on inferences of basic
stellar parameters to varying degrees.
Empirically-derived and reliable estimates are of particular value for
increasing our confidence in the end-product results built upon these inputs.

Basic stellar parameters such as effective temperature and surface gravity
can be inferred using one (or more) of several types of observations, e.g.
spectroscopy, photometry, interferometry, etc.
This inference can be performed using theoretical models or by building an
empirical calibration library.
For example, an observed stellar spectrum could be matched against a library of
theoretical spectra generated using stellar atmosphere models, or against a
library of observed spectra of ``standard stars'', serving as calibrators.
%Fundamental stellar parameters, such as mass and radius, which cannot be
%directly inferred from the available observations can often be inferred by
%invoking theoretical stellar models, e.g. fitting model isochrones to
%spectroscopic measurements.
Regardless of the approach, be it theoretical or empirical, the methods used
for the inference of stellar parameters are traditionally ``deterministic''.
A deterministic model can be loosely described as one where a particular
observational input always returns a single-valued output for a parameter of
interest.
In this framework, nature itself has no variance and our underlying model is
considered to be a perfect description of reality.

An alternative approach for inferring model parameters is a ``probabilistic''
one.
Unlike the deterministic case, a single observational input is intepreted to be
caused by a range of possible model parameters, described by a probability
distribution.
This statement is true even in the case of a perfect observation of infinite
signal-to-noise, since our underlying model itself comes with uncertainty.
In recent years there has been a shift towards such methods in several areas of
 astronomy and particularly within the exoplanet community.
For example, \citet{wolfgang:2015} considere the mass-radius relationship
of exoplanets to be probabilistic, since a planet of a particular radius could
have a range of masses thanks to variances in compositions, environment and
other circumstances.
These recent demonstrations in exoplanetary science prompt us to consider the
need for treating the parent stars in the same probabilistic framework, with
potential applications spanning many fields of astronomy.

The demand for probilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, it also has a practical purpose.
When using data published in the astronomical literature to, for example, infer
relationships between observed parameters, that inference can be performed as
the final stage in a hierarchical treatment.
This is useful if one wants to account for multi-dimensional uncertainties by
marginalizing over the `true' parameter values upon which the noisy
observations are dependent.
This point is discussed further in \textsection\ref{sec:HBM}.

### 2015-06-05 version1.0/manuscript.tex
adding words

\altaffiltext{1}{Dept. of Physics, University of Oxford, UK}

@@

TBD

>>>>

\altaffiltext{1}{Dept. of Physics, University of Oxford, UK; email:
		 ruth.angus@astro.ox.ac.uk}

@@

% aims
The relations between observable stellar parameters are usually assumed to be
deterministic, that is, given an infinitely precise measurement of independant
variable, `$x$', and some model, the value of dependant variable, `y' can be
known exactly.
In practise this assumption is rarely valid and intrinsic stochasticity means
that two stars with exactly the same `$x$', will have slightly different
`$y$`s.
% methods
The relation between short-timescale brightness fluctuations (flicker) of stars
and both stellar density and surface gravity \citep{bastien:2013} are two such
stochastic relations.
We recalibrate these relations in a probabilistic framework, using
Hierarchical Bayesian Modelling (HBM) and constrain the instrinsic scatter in
the relations.
Our treatment also fully accounts for the observational uncertainties on all
variables.
% results
These new relations provide probability distributions over values of $\log(g)$
or $\rho_*$ for given measurements of flicker.
Such probability distributions are both more representative of what we know
about stars and will be useful for subsequent studies using probabilistic
inference.
